---
title: "Practical Machine Learning Project"
date: "23/08/2021"
output: html_document
---

In this project, the data from accelerometers is divided into two sets: training and testing. The goal of the project is to predict the manner in which they did the exercise.
```{r}
# Reading the file
training <- read.csv("G:/Course/PracticalMachineLearning/pml-training.csv")
dim(training)
testing <- read.csv("G:/Course/PracticalMachineLearning/pml-testing.csv")
dim(testing)
```

Loading the required libraries
```{r, message=FALSE, warning=FALSE}
library(caret)
library(kernlab)
library(randomForest)
library(rattle)
library(lattice)
```
Cleaning the dataset
```{r}
# Removing NA Value Columns
training<- training[, colSums(is.na(training)) == 0]
testing<- testing[, colSums(is.na(testing)) == 0]
# Remmoving unwanted columns
training <- training[,-c(1:7)]
testing <- testing[,-c(1:7)]
# Removing zero variance columnbs
nvz <- nearZeroVar(training)
training <- training[,-nvz]
dim(training)
dim(testing)
```

```{r}
# Dividing training data into train and test to build the model
t <- createDataPartition(training$classe, p=0.7, list = FALSE)
trainData <- training[t,]
trainTest <- training[-t,]
# Defining control to be used with  models
ctrl <- trainControl(method = "cv", number=3 )
```
Predicting with trees
```{r}
m_trees <- train(classe~., method="rpart", data=trainData, trControl=ctrl)
p_trees <- predict(m_trees, trainTest)
confusionMatrix(p_trees, factor(trainTest$classe))
plot(m_trees)
```
Predicting with Support Vector Machine
```{r}
m_svm <- train(classe~., method="svmLinear", data=trainData, trControl = ctrl, tuneLength = 5)
p_svm <- predict(m_svm, trainTest)
confusionMatrix(p_svm, factor(trainTest$classe))
# plot(m_svm)
```
Predicting with Random Forest
```{r}
m_rf <- train(classe~., method="rf", data=trainData, trControl=ctrl)
p_rf <- predict(m_rf, trainTest)
confusionMatrix(p_rf, factor(trainTest$classe))
plot(m_rf)
```
Predicting with 
```{r, results='hide'}
m_gbm <- train(classe~., data=trainData, method="gbm", trControl = ctrl)
```
```{r}
p_gbm <- predict(m_gbm, trainTest)
confusionMatrix(p_gbm, factor(trainTest$classe))
plot(m_gbm)
```

## Result
From all the above models, we can see that the best prediction is given by Random Forest.
Accuracy = 0.9924 and Out of Sample Error = 0.0076

Using Random Forest Model to predict the values on test dataset as provided
```{r}
p <- predict(m_rf, testing)
print(p)
```
